---
title: "JSCBB Fume Hood Contest Write-Up Code"
author: "Ziyu Li"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(rstatix)
library(ggpubr)
```

# Data Cleaning
```{r The-Data-01, echo=FALSE}
file <- read.delim("JSCBB_Contest_2022_Hood_Data.txt", 
                   header = TRUE, sep = "\t")
#separate them into sash height vs. slider gap files
sash_height <- select(file, Hood.Number, Floor, Wing, SashP1, SashP2, Sash1, Sash2,
                      Sash3, Sash4, Sash5, Sash6, Sash7, Sash8, Sash9)
slider_gap <- select(file, Hood.Number, Floor, Wing, SliderP1, SliderP2, Slider1,
                     Slider2, Slider3, Slider4, Slider5, Slider6, Slider7,
                     Slider8, Slider9)
#removing NA's from the dataset. 
sash_height <- na.omit(sash_height)
slider_gap <- na.omit(slider_gap)
#and making sure all the measured values are "Numeric"
sash_height[, 4:ncol(slider_gap)] <- lapply(4:ncol(sash_height), function(x) as.numeric(sash_height[[x]]))
slider_gap[, 4:ncol(slider_gap)] <- lapply(4:ncol(slider_gap), function(x) as.numeric(slider_gap[[x]]))

#Door information
door <- read.delim("JSCBB_Contest_2022_Door_Data.txt", 
                   header = TRUE, sep = "\t")
door <- na.omit(door) #removing NAs, there were no NAs in the dataset
Pdoor <- c()
for (i in 1:10){
  Pdoor[i] <- sum(door[,i+3] == 1)/nrow(door)
}


```

After clearing the data and removing all the comments, we end up with 92 fume hood for vertical sash height and 75 fume hoods for horizontal slider gap with complete and consistent data throughout the contest.

Graphing some things for the paper:
```{r}
lineShape <- c(1,2,3)
par(mar = c(5,4.5,4,4.5))
plot(x = seq(0,10, by = 1), y = colMeans(sash_height[,4:14]), type = 'b', lwd = 3, 
     col = "#80cbc4", las = 1, lty = lineShape[1],
     main = "Average Fume Hood and Door Behavior \nwith Each Data Collection",
     ylab = "Average Opening [in]", xlab = "Data Collection Point", 
     ylim = c(0, 5.5), xaxt = 'n',cex.main = 1.2, cex.lab = 1,
     cex.axis = 1)
lines(x = seq(0,10, by = 1), y = colMeans(slider_gap[,4:14]), type = 'b', 
      lty = lineShape[2], lwd = 3, col = "#f85f8f")

par(new = TRUE)
plot(x = seq(0,10, by = 1), y = c(NA,Pdoor), type = 'b', lwd = 3, 
     col = "#a299d4", main = "", ylab = "", xlab = "", axes = FALSE,
     ylim = c(0.5, 1), xaxt = 'n',cex.main = 1.2, cex.lab = 1, 
     cex.axis = 1, lty = lineShape[3])
axis(side = 4, las = 1)
mtext("Proportion", side = 4, line = 3, las = 0)
axis(side = 1, at = seq(0,10, by = 1), 
     labels = c("P01", "P02", "C01", "C02", "C03", "C04", "C05", "C06", 
                "C07", "C08", "C09"), cex.axis = 1)
legend("bottomleft", legend = c("Proportion Closed Doors",
                                "Horizontal Slider Total Gap", "Vertical Sash Height"), 
       lty = rev(lineShape), cex = 0.85, col = c("#a299d4","#f85f8f", "#80cbc4"), 
       bty = "n", lwd = 3 )
```
# Assumptions
Outliers removed and examine normality with Shapiro test due to small sample size. 
```{R}
#convert to long-format
sash_height_long <- sash_height %>%
  gather(key = "collection", value = "height", SashP1, SashP2, Sash1, Sash2, 
         Sash3, Sash4, Sash5, Sash6, Sash7, Sash8, Sash9) %>%
  convert_as_factor(Hood.Number, collection)
#first, get rid of labs that did not change, since some hoods aren't in use much
summary <- sash_height_long %>%
  group_by(Hood.Number) %>%
  get_summary_stats(height)
chosen <- summary$Hood.Number[summary$mean == 0]
chose <- as.factor(chosen)
sash_height_long <- sash_height_long[! sash_height_long$Hood.Number %in% chose,]

# Identify outliers
outliers_sash <- sash_height_long %>%
  group_by(collection) %>%
  identify_outliers(height)
outliers_sash <- as.data.frame(outliers_sash)

# prepare for removal
chosen_sash <- matrix(data = NA, nrow = nrow(outliers_sash), ncol = 1)
for (i in 1:dim(outliers_sash)[1]){
  if (outliers_sash$is.extreme[i] == TRUE){
    chosen_sash[i] <- as.character(outliers_sash[i,2])
  } else {
    chosen_sash[i] <- NA
  }
}
chosen_sash <- na.omit(chosen_sash)
facchosen_sash <- factor(chosen_sash)
sash_height_clean <- sash_height_long[! sash_height_long$Hood.Number %in% facchosen_sash,]

#check within subject samples
summary_sash <- sash_height_clean %>%
  group_by(collection) %>%
  get_summary_stats(height)
data.frame(summary_sash)
#Shapiro-Wilks. Null = normality is met, alternative = not met
shapiroTest_sash <- sash_height_clean %>%
  group_by(collection) %>%
  shapiro_test(height)
shapiroTest_sash
```
```{r}
range(shapiroTest_sash$p)
```
There are 46 between-subject samples. So 46 samples per time slot for sashes. P-values range from $1.37 \times 10^{-11}$ to $1.23 \times 10^{-7}$ for vertical sash height. Indicating non-normality. 

```{r}
#converting to long form and identify outliers
slider_gap_long <- slider_gap %>%
  gather(key = "collection", value = "gap", SliderP1, SliderP2, Slider1, Slider2, 
         Slider3, Slider4, Slider5, Slider6, Slider7, Slider8, Slider9) %>%
  convert_as_factor(Hood.Number, collection)
#get rid of the hoods that weren't used
summary_slider <- slider_gap_long %>%
  group_by(Hood.Number) %>%
  get_summary_stats(gap)
chosen_slider <- summary$Hood.Number[summary$mean == 0]
chose_slider <- as.factor(chosen_slider)
slider_gap_long <- slider_gap_long[! slider_gap_long$Hood.Number %in% chose_slider,]

outliers_slider <- slider_gap_long %>%
  group_by(collection) %>%
  identify_outliers(gap)
outliers_slider <- as.data.frame(outliers_slider)

#Clean out the influential outliers
chosen_slider <- matrix(data = NA, nrow = nrow(outliers_slider), ncol = 1)
for (i in 1:dim(outliers_slider)[1]){
  if (outliers_slider$is.extreme[i] == TRUE){
    chosen_slider[i] <- as.character(outliers_slider[i,2])
  } else {
    chosen_slider[i] <- NA
  }
}
chosen_slider <- na.omit(chosen_slider)
facchosen_slider <- factor(chosen_slider)
slider_gap_clean <- slider_gap_long[! slider_gap_long$Hood.Number %in% facchosen_slider,]

#within sample n
summary_slider <- slider_gap_clean  %>%
  group_by(collection) %>%
  get_summary_stats(gap)
data.frame(summary_slider)
#Shapiro-Wilks. Null = normality is met, alternative = not met
shapiroTest_slider <- slider_gap_clean  %>%
  group_by(collection) %>%
  shapiro_test(gap)
shapiroTest_slider
```

```{R}
range(shapiroTest_slider$p)
```

There are 28 within-subject samples. P-value for the Shapiro-Wilks test is from $4.74 \times 10^{-10}$ to $ 6.78 \times 10^{-6}$ for horizontal sliders. 

Sphericity is not needed since the correction is built in. 

# Longitudinal ANOVA Test
```{r}
vert.anov <- anova_test(data = sash_height_clean, dv = height, wid = Hood.Number, within = collection)
vert.anov
get_anova_table(vert.anov)
```



```{R}
horiz.anov <- anova_test(data = slider_gap_clean, dv = gap, wid = Hood.Number, within = collection)
horiz.anov
get_anova_table(horiz.anov)
```


# Post-Hoc
```{r}
order <- c("SashP1", "SashP2", "Sash1", "Sash2", "Sash3", "Sash4", 
           "Sash5", "Sash6", "Sash7", "Sash8", "Sash9")
pwc <- sash_height_clean %>%
  pairwise_t_test(
    height ~ collection, paired = TRUE,
    p.adjust.method = "BH"
    )

pwc <- pwc %>% add_xy_position(x = "collection")
ymax <- 10
#clean up plotting
bxp <- ggboxplot(sash_height_clean, x = "collection", 
                 y = "height", outlier.shape = NA ,  
                 order = order, ylim = c(0, ymax),
                 xlab = "Data Collection Point",
                 ylab = "Height [in]",
                 main = "Vertical Sash Data with Results of Pairwise t-Test") +
    theme(
      plot.title = element_text(hjust = 0.5)
      )
pwc[pwc$p.adj.signif == 'ns',] <- NA
pwc_new <- na.omit(pwc)
pwc_new
new_names <- c("P01", "P02", "C01", "C02", "C03", "C04", "C05", 
               "C06", "C07", "C08","C09")

# Update the levels of the x-axis variable in the plot
bxp <- bxp +
  scale_x_discrete(labels = new_names)

# The rest of your code
min <- 0
max <- 0
for (i in 1:nrow(pwc_new)) {
  for (j in 1:length(order)) {
    if (pwc_new$group1[i] == order[j]) {
      min <- j
    }
    if (pwc_new$group2[i] == order[j]) {
      max <- j
    }
  }
  if (min < max) {
    pwc_new$xmin[i] <- min
    pwc_new$xmax[i] <- max
  } else {
    pwc_new$xmax[i] <- min
    pwc_new$xmin[i] <- max
  }
}
pwc_new$y.position <- ymax - 6 + seq(1:nrow(pwc_new)) * 0.8
bxp + 
  stat_pvalue_manual(pwc_new) +
  labs(
    subtitle = get_test_label(vert.anov, detailed = TRUE),
    caption = get_pwc_label(pwc_new)
  )
```
There was no significant pairs in sliders. 

# Energy
## Bootstrap CI
```{r}
set.seed(123)
# Vertical Sash Before
sashBeforeRaw <- sash_height$SashP1
N <- 2000 # num of bootstrap samples
n <- length(sashBeforeRaw) # number of draws within each bootstrap sample
sashBeforeBoot <- c() # empty vector for bootstrap means
for (i in 1:N) {
  bootSample <- sample(sashBeforeRaw, size = , replace = TRUE)
  sashBeforeBoot[i] <- mean(bootSample)
}

sashAfterRaw <- sash_height$Sash9
sashAfterBoot <- c() #empty vector for bootstrap means
for (i in 1:N) {
  bootSample <- sample(sashAfterRaw, size = n, replace = TRUE)
  sashAfterBoot[i] <- mean(bootSample)
}

par(mfrow = c(2,2))
hist(sash_height$SashP1)
hist(sash_height$Sash9)
hist(sashBeforeBoot)
hist(sashAfterBoot)
```
There are still slight skewness but bootstrap method has moslty corrected non-normality. 
```{r}
par(mfrow = c(2, 2))
hist(sashBeforeRaw, freq = FALSE, col = "#80cbc4", xlim = c(0,20), ylim = c(0,0.9),
     main = "Raw Data Before Contest", xlab = "Vertical Sash Height [in]",
     breaks = 15)
hist(sashBeforeBoot, freq = FALSE, col = "#80cbc4", xlim = c(0, 7), ylim = c(0, 0.9),
     main = "Bootstrap Data Before Contest", xlab = "Mean Vertical Sash Height [in]",
     breaks = 15)
hist(sashAfterRaw, freq = FALSE, col = "#f85f8f", xlim = c(0,20), ylim = c(0,0.9),
    main = "Raw Data End of Contest", xlab = "Vertical Sash Height [in]",
    breaks = 15)
hist(sashAfterBoot, freq = FALSE, col = "#f85f8f",xlim = c(0, 7), ylim = c(0, 0.9),
     main = "Bootstrap Data End of Contest", xlab = "Mean Vertical Sash Height [in]",
     breaks = 10)
```
Means of raw
```{r}
averageBefore <- mean(sashBeforeBoot); averageBefore
averageAfter <- mean(sashAfterBoot); averageAfter
```

Empirical confidence intervals:
```{r}
sashBeforeCI <- quantile(sashBeforeBoot, c(0.025, 0.975))
sashAfterCI <- quantile(sashAfterBoot, c(0.025, 0.975))
sashBeforeCI
sashAfterCI
```
## Face velocity
```{r}
velocities <- read.delim("BIOT_Fume_Hood_Velocities.txt", 
                     header = TRUE, sep = "\t")
names(velocities) <- c("Room", "Manufacturer", "Model", "AuditDate", "Lfpm", "Cfpm", "Rfpm","Avgfpm")
velocities <- na.omit(velocities)
hist(velocities$Avgfpm, breaks = 20) #since this is normal, we can just use the mean for sash height of 14
length(velocities$Avgfpm)
```
Approx linear, so we use average. 
```{r}
w <- 6
velocity_closed <- 100/(w * (2/13)); velocity_closed # not important
velocity_14 <- mean(as.numeric(velocities$Avgfpm)); velocity_14
```

```{R}
flow14 <- velocity_14 * w * 14/12; flow14 #flow rate at 14 inches of sash height
```

## Flowrate
interpolation
```{r}
heightIn <- c(2, 14)
flow <- c(100, flow14) # the "y"
flowBeforeAvg <- approx(heightIn, flow, xout = averageBefore)$y
flowBeforeAvg #avg flow rate before is 191 cfm
flowAfterAvg <- 100
# 95% CI for before contest
flowBeforeCI <- approx(heightIn, flow, xout = sashBeforeCI)$y
flowBeforeCI # give (135, 253)

# 95% CI for after contest
flowAfterCI <- approx(heightIn, flow, xout = sashAfterCI)$y
#flowAfterCI will give NA for lower bound due to less than 2in
flowAfterCI[1] <- 100
flowAfterCI #will give (100, 119)

```

## Power & Energy Cooling/Heating
Now power of heating and cooling:
$BTUH_{cooling/heating} = CFM *\Delta T_{cooling/heating} * 0.89$, where $\Delta T_{heating} = 17$ and $\Delta T_{cooling} = 23$ . Multiply that by heating (5125) & cooling degree days (1586), we can get the BTU for cooling and heating for the whole year.
```{r}
# Power Heating
## Before
powerHeatingBeforeAvg <- 0.89 * flowBeforeAvg * 17; powerHeatingBeforeAvg
powerHeatingBeforeCI <- 0.89 * flowBeforeCI * 17; powerHeatingBeforeCI
## After
powerHeatingAfterAvg <- 0.89 * flowAfterAvg * 17; powerHeatingAfterAvg
powerHeatingAfterCI <- 0.89 * flowAfterCI * 17; powerHeatingAfterCI
```
Mean heating before, after: 2886, 1513
CI before, after: (2038, 3827), (1513, 1807)

```{r}
# Power Cooling
## Before
powerCoolingBeforeAvg <- 0.89 * flowBeforeAvg * 23; powerCoolingBeforeAvg
powerCoolingBeforeCI <- 0.89 * flowBeforeCI * 23; powerCoolingBeforeCI
## After
powerCoolingAfterAvg <- 0.89 * flowAfterAvg * 23; powerCoolingAfterAvg
powerCoolingAfterCI <- 0.89 * flowAfterCI * 23; powerCoolingAfterCI
```
Mean cooling before, after: 3906, 2047
CI before, after: (2758, 5178), (2047, 2445)

```{r}
# Energy Heating
## Before
EHeatingBeforeAvg <- 5125 * powerHeatingBeforeAvg
EHeatingBeforeAvg; sprintf("%.2e", EHeatingBeforeAvg)
EHeatingBeforeCI <- 5125 * powerHeatingBeforeCI
sprintf("%.2e", EHeatingBeforeCI)

EHeatingAfterAvg <- 5125 * powerHeatingAfterAvg
EHeatingAfterAvg
EHeatingAfterCI <- 5125 * powerHeatingAfterCI
sprintf("%.2e", EHeatingAfterCI)
```
Mean heating before, after: 1.48e+07, 7.75e+06
CI before, after: (1.04e+07, 1.96e+07), (7.75e+06, 9.26e+06)

```{r}
# Energy Cooling
## Before
ECoolingBeforeAvg <- 1586 * powerCoolingBeforeAvg
ECoolingBeforeAvg; sprintf("%.2e", ECoolingBeforeAvg)
ECoolingBeforeCI <- 1586 * powerCoolingBeforeCI
sprintf("%.2e", ECoolingBeforeCI)

ECoolingAfterAvg <- 1586 * powerCoolingAfterAvg
ECoolingAfterAvg
ECoolingAfterCI <- 1586 * powerCoolingAfterCI
sprintf("%.2e", ECoolingAfterCI)
```
Mean heating before, after: 6.19e+06, 3.25e+06
CI before, after: (4.37e+06, 8.21e+06), (3.25e+06, 3.88e+06)

## Total Energy
1kWh = 3412.14 BTU
```{r}
# Total energy in kWh before
totEBeforeAvg <- (EHeatingBeforeAvg + ECoolingBeforeAvg)/3412.14
totEBeforeAvg
totEBeforeCI <- (EHeatingBeforeCI + ECoolingBeforeCI)/3412.14
totEBeforeCI
```
Total energy before per hood per year is 6141 kWh, with CI (4344.418,8155.074)
```{r}
# Total energy in kWh after
totEAfterAvg <- (EHeatingAfterAvg + ECoolingAfterAvg)/3412.14
totEAfterAvg
totEAfterCI <- (EHeatingAfterCI + ECoolingAfterCI)/3412.14
totEAfterCI
```
Total energy after per hood per year is 3224 kWh with CI (3224, 3850.719)
```{r}
# Energy savings in kWh
totESaveAvg <- totEBeforeAvg - totEAfterAvg
totESaveAvg
totESaveCI <- totEBeforeCI - totEAfterCI
totESaveCI
```
Total energy savings is 2927.364 kWh per hood per year with CI (1120.439, 4304.355)

## Cost
```{r}
# cost for electricity
costElectricity <- 0.1234 #$/kWh
costSteamPound <- 27.418 #$/klb
costSteam <- costSteamPound*3412.14/(1000*1194)
costSteam
```
cost of steam is $0.07835/kWh.
```{r}
# Energy savings in kWh for heating
diffEHeatingAvg <- (EHeatingBeforeAvg - EHeatingAfterAvg)/3412.14
diffEHeatingAvg
diffEHeatingCI <- (EHeatingBeforeCI - EHeatingAfterCI)/3412.14
diffEHeatingCI 
# Energy savings in kWh for cooling
diffECoolingAvg <- (ECoolingBeforeAvg - ECoolingAfterAvg)/3412.14
diffECoolingAvg
diffECoolingCI <- (ECoolingBeforeCI - ECoolingAfterCI)/3412.14
diffECoolingCI
```
Energy savings for heating is 2063.433 kWh per hood per year, with CI (789.7726, 3034.0436)
That for cooling is 863.9302 kWh, with CI (330.6666, 1270.3110)


```{r}
# Costs Heating
costSaveHeatingAvg <- diffEHeatingAvg*costSteam
costSaveHeatingAvg
costSaveHeatingCI <- diffEHeatingCI*costSteam
costSaveHeatingCI
# Cost Cooling
costSaveCoolingAvg <- diffECoolingAvg*costElectricity
costSaveCoolingAvg
costSaveCoolingCI <- diffECoolingCI*costElectricity
costSaveCoolingCI
```
Cost for heating is $162 per hood per year with CI (61.88143, 237.72788)
That for cooling is $107, with CI (40.80425, 156.75637)

```{r}
# Total cost savings
costSaveAvg <- costSaveHeatingAvg + costSaveCoolingAvg
costSaveAvg
costSaveCI <- costSaveHeatingCI + costSaveCoolingCI
costSaveCI
```
Cost is $268 per hood per year with CI (102.6857, 394.4842)

## Building Wide Cost and Energy
```{r}
# building energy savings
buildingESaveAvg <- totESaveAvg*170
buildingESaveAvg
buildingESaveCI <- totESaveCI*170
buildingESaveCI

# building cost saving
buildingCostSaveAvg <- costSaveAvg*170
buildingCostSaveAvg
buildingCostSaveCI <- costSaveCI*170
buildingCostSaveCI
```
With 170 fume hoods, the energy savings is 497651.8 kWh, CI (190474.7,731740.3).
And cost is $45608.65, CI (17456.57, 67062.32)

# Conclusion
```{r}
# average reduction in sash height %
100*(averageBefore - averageAfter)/averageBefore
# average reduction in energy
100*(totEBeforeAvg - totEAfterAvg)/totEBeforeAvg
```